---
title: "Create_Dataframe"
format: html
editor_options: 
  chunk_output_type: console
---


This markdown loads the necessary raster datasets - transforms them to match the same spatial extent and coordinate system as the fractional cover data, and converts to data.table format and merges into a single dataframe representing all raster variables.

This table is used for GAMs in section 1 and 2 of the paper.

```{r}
#| label: setup
pacman::p_load(tidyverse, lubridate, terra, sf, fields, arrow, data.table)
outputFolder <- "./outputs/tables"

# Function to process rainfall datasets. Pivot longer to have date column.
processRainfall <- function(path, dataset_name) {
  precip <- rast(path)
  dates <- time(precip)
  precip <- precip %>% 
    project(blankRast) %>%
    as.data.frame(xy = T, cells = T, na.rm = F)

  names(precip)[4:ncol(precip)] <- as.character(dates)

  precip <- precip %>%  
    pivot_longer(., cols = 4:ncol(.), names_to = "date", values_to = dataset_name)
  setDT(precip)

  return(precip)
}
```


# Create template raster for reprojection

```{r}
path_pv <- "./outputs/raster/guerschman_fc_monthly/fc_monthly_pv_20010101_20230101.nc"

# Create blank raster - used to project all datasets to same coordinate system
blankRast <- rast(path_pv)[[1]]
blankRast[!is.na(blankRast)] <- NA
names(blankRast) <- "blankRast"

# Polygon extent of study area
extent <- read_sf("./outputs/vector/bonbon_buffer.gpkg") %>% 
  st_geometry()

# Get Dates from pv file names
dates <- rast(path_pv) %>% 
  time()
```

# Load FC variables

```{r}
path_npv <- "./outputs/raster/guerschman_fc_monthly/fc_monthly_npv_20010101_20230101.nc"
path_bare <- "./outputs/raster/guerschman_fc_monthly/fc_monthly_bare_20010101_20230101.nc"

##########
### PV ###
##########

pv <- rast(path_pv)
pv <- as.data.frame(pv, xy = T, cells = T, na.rm = F)
names(pv)[4:ncol(pv)] <- as.character(dates)

pv <- pv %>% 
  pivot_longer(cols = 4:ncol(.), names_to = "date", values_to = "pv")
setDT(pv)

###########
### NPV ###
###########

npv <- rast(path_npv)
npv <- as.data.frame(npv, xy = T, cells = T, na.rm = F)
names(npv)[4:ncol(npv)] <- as.character(dates)

npv <- npv %>% 
  pivot_longer(cols = 4:ncol(.), names_to = "date", values_to = "npv")
setDT(npv)

############
### Bare ###
############

bare <- rast(path_bare)
bare <- as.data.frame(bare, xy = T, cells = T, na.rm = F)
names(bare)[4:ncol(bare)] <- as.character(dates)

bare <- bare %>% 
  pivot_longer(cols = 4:ncol(.), names_to = "date", values_to = "bare")
setDT(bare)

#############################
### Bind cover components ###
#############################

df <- cbind(pv, npv[,5]) %>% 
  cbind(., bare[,5])

rm(list = c("pv", "npv", "bare"))
gc(full = T, reset = T)
```

# AGCD Precipitation

```{r eval = F}
# p_agcd <- processRainfall("./outputs/raster/agcd_precipitation/agcd_v2-0-1_precip_total_r001_monthly_2000_2022.nc", "p_agcd")
# 
# setkeyv(p_agcd, c("date", "x","y"))
# df <- merge(df, 
#             p_agcd %>% select(-cell), 
#             by = c("date", "x", "y"),
#             all.x = T)
# 
# rm(p_agcd)
# gc(full = T, reset = T)
# 
# p_agcd_3 <- processRainfall("./outputs/raster/agcd_precipitation/rolling/agcd_v2-0-1_precip_total_r001_monthly_2000_2022.nc_3mo_sum.nc", "p_agcd_3")
# 
# setkeyv(p_agcd_3, c("date", "x","y"))
# df <- merge(df, 
#             p_agcd_3 %>% select(-cell), 
#             by = c("date", "x", "y"),
#             all.x = T)
# 
# rm(p_agcd_3)
# gc(full = T, reset = T)
# 
# p_agcd_12 <- processRainfall("./outputs/raster/agcd_precipitation/rolling/agcd_v2-0-1_precip_total_r001_monthly_2000_2022.nc_12mo_sum.nc", "p_agcd_12")
# 
# setkeyv(p_agcd_12, c("date", "x","y"))
# df <- merge(df, 
#             p_agcd_12 %>% select(-cell), 
#             by = c("date", "x", "y"),
#             all.x = T)
# 
# rm(p_agcd_12)
# gc(full = T, reset = T)

```

# Statistical zones

Vector datasets including station boundaries, etc. used for grouping in modelling Converted to raster data

```{r}
################
### Stations ###
################

stations <- read_sf("./outputs/vector/pastoral_stations.gpkg") %>%
  vect() %>% 
  rasterize(blankRast, field = "NAME") %>% 
  as.data.frame(xy = T, cells = T, na.rm = F)
names(stations)[4] <- "station"
setDT(stations)

setkeyv(stations, c("x","y"))

df <- merge(df, 
            stations %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(stations)
gc(full = T, reset = T)


#######################
### IBRA SubRegions ###
#######################

ibra_sr <- read_sf("./outputs/vector/ibra_subregions.gpkg") %>% 
  st_transform("EPSG:28353") %>% 
  vect() %>% 
  rasterize(blankRast, field = "SUB_NAME_7") %>% 
  as.data.frame(xy = T, cells = T)
setDT(ibra_sr)

setkeyv(ibra_sr, c("x","y"))

df <- merge(df, 
            ibra_sr %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(ibra_sr)
gc(full = T, reset = T)

###################
### Landsystems ###
###################

l_sys <- read_sf("./outputs/vector/pastoral_landsystems.gpkg") %>% 
  st_transform("EPSG:28353") %>% 
  vect() %>% 
  rasterize(blankRast, field = "LANDSYSTEM") %>% 
  as.data.frame(xy = T, cells = T)
setDT(l_sys)

setkeyv(l_sys, c("x","y"))

df <- merge(df, 
            l_sys %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(l_sys)
gc(full = T, reset = T)

l_sys_id <- read_sf("./outputs/vector/pastoral_landsystems.gpkg") %>% 
  st_transform("EPSG:28353") %>% 
  vect() %>% 
  rasterize(blankRast, field = "OBJECTID") %>% 
  as.data.frame(xy = T, cells = T)
setDT(l_sys_id)

setkeyv(l_sys_id, c("x","y"))

df <- merge(df, 
            l_sys_id %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(l_sys_id)
gc(full = T, reset = T)

######################
### NVIS SubGroups ###
######################

nvis_s <- rast("./data/raster/NVIS6_0_AUST_EXT_MVS_ALB.tif") %>% 
  project(blankRast)
nvis_s[nvis_s == "Unknown/No data"] <- NA
nvis_s <- nvis_s %>% 
  as.data.frame(xy = T, cells = T)
setDT(nvis_s)

setkeyv(nvis_s, c("x","y"))

df <- merge(df, 
            nvis_s %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(nvis_s)
gc(full = T, reset = T)

###################
### NVIS Groups ###
###################

nvis <- rast("./data/raster/NVIS6_0_AUST_EXT_MVG_ALB.tif") %>% 
  project(blankRast)
nvis[nvis == "Unknown/no data"] <- NA
nvis <- nvis %>% 
  as.data.frame(xy = T, cells = T)
setDT(nvis)

setkeyv(nvis, c("x","y"))

df <- merge(df, 
            nvis %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(nvis)
gc(full = T, reset = T)

######################
### Nearby Bon Bon ###
######################

zone <- read_sf("./outputs/vector/bonbon_buffer.gpkg") %>% 
  vect() %>% 
  rasterize(blankRast, field = "zone")
zone[is.na(zone)] <- 3
zone <- zone %>% 
  as.data.frame(xy = T, cells = T)

setDT(zone)
setkeyv(zone, c("x","y"))
df <- merge(df, 
            zone %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(zone)
gc(full = T, reset = T)

############################
### Station Fence Buffer ###
############################


fences <- read_sf("./outputs/vector/fences.gpkg") %>%
  rasterize(blankRast, field = "fence") %>% 
  as.data.frame(xy = T, cells = T)
setDT(fences)

setkeyv(fences, c("x","y"))

df <- merge(df, 
            fences %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(fences)
gc(full = T, reset = T)

#################
### Dog Fence ###
#################

dog_fence <- read_sf("./outputs/vector/dog_fence.gpkg") %>% 
  rasterize(blankRast, field = "inside") %>% 
  as.data.frame(xy = T, cells = T, na.rm = F)
setDT(dog_fence)

setkeyv(dog_fence, c("x","y"))

df <- merge(df, 
            dog_fence %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(dog_fence)
gc(full = T, reset = T)
```

# Water areas

```{r}
water <- read_sf("./outputs/vector/water_courses.gpkg") %>% 
  add_column(water = 1) %>% 
  vect() %>% 
  rasterize(blankRast, field = "water")

water[water != 1] <- NA

water <- water %>% 
  as.data.frame(xy = T, cells = T)
setDT(water)

setkeyv(water, c("x","y"))

df <- merge(df,
            water %>% select(-cell), 
            by = c("x", "y"),
            all.x = T)

rm(water)
gc(full = T, reset = T)
```

# Mask

```{r}
mask <- rast("./outputs/raster/mask.tif")
names(mask) <- "mask"
mask <- mask %>% 
  as.data.frame(xy = T, cells = T)
setDT(mask)
setkeyv(mask, c("x", "y"))

df <- merge(df, mask %>% select(-cell),
            by = c("x", "y"),
            all.x = T)
rm(mask)
gc(full = T, reset = T)

df <- df %>%
  filter(mask != 1)

df <- df %>%
  select(-mask)
```

```{r}
df <- df %>% 
  mutate(year = as.factor(year(df$date)),
         .after = 4)
```


```{r}
names(df) <- c("x", "y", "cell", "date", "year", "pv", "npv", "bare", "station", "ibra_subregion", "land_system", "land_system_id", "nvis_subgroup", "nvis_group", "study_zone", "fence_buffer", "dog_fence", "water")
```


# Save df

```{r}
arrow::write_parquet(df, file.path(outputFolder, "full_dataframe.parquet"))

# df %>%
#   .[sample(1:nrow(.), 10000),] %>%
#   st_as_sf(coords = c("x", "y")) %>% st_geometry() %>%
#   plot()
```
